Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-11 Methods-and-Results.tex
shortened + typos + refs

After receiving ethical approval from the University Masked for Peer
Review, we recruited 23 non-Dutch migrants using the local paid
participant pool. Participants reported on their interactions for at
least 30 days with two daily measures (capturing the morning and
afternoon). With this design, we aimed at getting 50-60 measurements per
participant (\textit{M} = 53.26, \textit{SD} = 16.72, \textit{total N} =
1,225). This is a common number of measurements found in experience
sampling studies and offers sufficient power to model processes within
and between participants
\citep[e.g., for a systematic review see][]{AanhetRot2012}. Participants
were compensated for their participation with up to 34 Euros -- each two
Euros for pre- and post-questionnaire and 50 Eurocents for every
experience sampling measurement. The sample consisted of relatively
young, educated, and western migrants from the global north (\(M_{age}\)
= 24.35, \(SD_{age}\) = 4.73, 19 women, 15 students). The sample
accurately describes the largest groups of migrants in the region
\citep[][]{GemeenteGroningen2015}.

@@

To test the prerequisite effect of intergroup contact, every experience
sampling measurement started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (In person interaction for at least 10 minutes)}''.
Our participants recorded between 2--51 interactions with Dutch outgroup
members (3.23--91.07\% of individual experience sampling measurements;
387 of all 1,225 experience sampling
responses)\footnote{Two participants only recorded two experience sampling measurements each and non of these included outgroup contacts. These participants are removed from any analyses that focus on outgroup contacts.}.

@@

We used a multilevel regression to check whether having an interaction
with an outgroup member had a situational (i.e., contemporaneous) effect
within the participants. We find that having an outgroup contact is
indeed associated with significantly more positive outgroup attitudes
(\textit{b} = 2.48, \textit{t}(1,200) = 4.37, \textit{p} \textless{}
.001, \textit{95\%CI}{[}1.37, 3.59{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently). Additionally, while multi-level regressions
are generally robust against unequal cell sizes, we correct for
inequalities by using centered predictors and reintroducing the means as
level two predictors (\citealp{Yaremych2021}; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.

@@

The main proposal of our article is that the success of an outgroup
contact might be explained by whether or not the contact fulfilled the
person's core situational need. This should, in turn, be due to a higher
perceived contact quality. We sequentially test whether the fulfillment
of the core need during an interaction is (1) related to more positive
outgroup attitudes, (2) higher perceived contact quality, and (3)
whether the variance explained by the core need is subsumed by the
perceived contact quality if considered jointly. We find that in the
multilevel models, the fulfillment of core situational needs during
outgroup contacts was associated with more positive outgroup attitudes
(random slopes model; \textit{b} = 0.17, \textit{t}(365) = 2.93,
\textit{p} = 0.004, \textit{95\%CI}{[}0.06, 0.29{]}) and also predicted
higher perceived contact quality (random intercept model; \textit{b} =
0.42, \textit{t}(365) = 8.07, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we consider the
influences of core need fulfillment and contact quality on outgroup
attitudes jointly, we find that virtually all variance is explained by
perceived contact quality (random slopes mode; \textit{b} = 0.23,
\textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}). We thus find support for our hypotheses and can conclude that
in this data set the fulfillment of core situational needs had a
significant influence on outgroup attitudes. Additionally, this effect
is likely explained by its effect through perceived contact quality.

@@

We tested the most general contact hypothesis as we did for Study 1. We
find that having an outgroup interaction is indeed associated with
significantly more positive outgroup attitudes within the participants
(random slopes model; \textit{b} = 2.83, \textit{t}(4,850) = 3.57,
\textit{p} \textless{} .001, \textit{95\%CI}{[}1.28, 4.38{]}), even
after controlling for having an interaction with a non-Dutch person
(which did not relate to outgroup attitudes independently). We again
added the participant means back into the model. We find that in this
data set participant-level outgroup contact proportions were also a
positive predictor of outgroup attitudes (\textit{b} = 26.55,
\textit{t}(110) = 2.90, \textit{p} = 0.004, \textit{95\%CI}{[}8.61,
44.46{]}). The relative number of non-outgroup interactions showed no
such effect (for full results see \tblref{tab:intergroupGeneralTblLong},
\fgrref{fig:ContactHypothesis}, as well as Online Supplementary Material
A). Thus, in our second data set we also find that outgroup contacts
show a positive effect on outgroup attitudes in the moment. We
additionally find an average between participant effect of the relative
number of interactions participants had.

@@

We again sequentially tested whether the fulfillment of the core need
during an interaction was (1) related to more positive outgroup
attitudes, (2) higher perceived interaction quality, and (3) whether the
variance explained by the core need is subsumed by the perceived
interaction quality if considered jointly. We find that in the
multilevel models, the fulfillment of core situational needs during
outgroup contacts was associated with more positive outgroup attitudes
(random slopes model; \textit{b} = 0.13, \textit{t}(826) = 4.18,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.07, 0.19{]}) and also
predicted higher perceived interaction quality (random slopes model;
\textit{b} = 0.40, \textit{t}(826) = 6.46, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.28, 0.52{]}). Additionally, if we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that virtually all variance is explained by
perceived interaction quality (random slopes mode; \textit{b} = 0.17,
\textit{t}(825) = 5.56, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.11, 0.23{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.03,
\textit{t}(825) = 1.47, \textit{p} = 0.141, \textit{95\%CI}{[}-0.01,
0.07{]}; also see \tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment} for full results). These results are
consistent with the results in Study 1. We, thus, find support for our
hypotheses that the fulfillment of core situational needs had a
significant influence on outgroup attitudes and that this effect is
likely explained by its effect through perceived interaction quality.

@@

We recruited 71 international medical students using contacts within the
University Medical School. We specifically targeted non-Dutch students,
who had recently arrived in the Netherlands. Participants reported on
their interactions for at least 30 days with two daily measures
(capturing the morning and afternoon). With this design, we aimed at
getting 50-60 measurements per participant (\textit{M} = 57.85,
\textit{SD} = 20.68, \textit{total N} = 4,107). As with the previous
studies, this offered sufficient power to model processes within
participants and will lend stronger weight to between-participant
results. Participants were compensated in the same manner as during
Study 1. The sample consisted of relatively young migrants, who were
mostly from the global north (\(M_{age}\) = 22.68, \(SD_{age}\) = 3.10,
59 women). The sample fairly accurately describes the local population
of young international medical professionals.

@@

In the multilevel regression, we find that having an outgroup
interaction was again associated with significantly more positive
outgroup attitudes within the participants (random slopes model;
\textit{b} = 5.57, \textit{t}(3,834) = 6.52, \textit{p} \textless{}
.001, \textit{95\%CI}{[}3.90, 7.23{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material A). Thus, in our third data set we
find that the within-person contemporaneous effect of intergroup contact
was consistent across all three studies.

@@

Analogous to the previous studies, we again sequentially tested whether
the fulfillment of the core need during an outgroup interaction was (1)
related to more positive outgroup attitudes, (2) higher perceived
interaction quality, and (3) whether the variance explained by the core
need is assumed by the perceived interaction quality if considered
jointly. We find that the fulfillment of the core need during outgroup
contacts was associated with more positive outgroup attitudes (random
slopes model; \textit{b} = 0.19, \textit{t}(1,601) = 5.29, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.12, 0.27{]}) and also predicted
higher perceived interaction quality (random slopes model; \textit{b} =
0.43, \textit{t}(1,605) = 8.35, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.33, 0.54{]}). Additionally, once we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that perceived interaction quality is a
substantially stronger predictor (random slopes mode; \textit{b} = 0.17,
\textit{t}(1,600) = 6.86, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.22{]}) and the unique variance explained by
core need fulfillment was roughly half of its original effect size
(\textit{b} = 0.11, \textit{t}(1,600) = 3.50, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.05, 0.17{]}; for full results see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). As with the previous two studies,
these results indicate that in this data set outgroup attitudes were
significantly predicted by the fulfillment of core situational needs and
the results suggest that this might be due to higher perceived
interaction quality.

@@

As we did with core need fulfillment, we sequentially tested whether the
fulfillment of Allport's contact conditions was (1) related to more
positive outgroup attitudes, (2) higher perceived interaction quality,
and (3) whether the variance explained by Allport's Conditions is
assumed by the perceived interaction quality if considered jointly. In
the multilevel models we find that the fulfillment of Allport's
Conditions during outgroup contacts was associated with more positive
outgroup attitudes (random slopes model; \textit{b} = 0.22,
\textit{t}(1,601) = 5.86, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.15, 0.29{]}) and also predicted higher perceived
interaction quality (random slopes model; \textit{b} = 0.65,
\textit{t}(1,605) = 11.82, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.54, 0.76{]}). Moreover, when we considered the
influences of Allport's Conditions and interaction quality on outgroup
attitudes jointly, we found that perceived interaction quality was a
substantially stronger predictor (random slopes mode; \textit{b} = 0.17,
\textit{t}(1,600) = 6.25, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.23{]}) and the unique variance explained by
Allport's Conditions was less than half of its original effect size
(\textit{b} = 0.09, \textit{t}(1,600) = 2.89, \textit{p} = 0.004,
\textit{95\%CI}{[}0.03, 0.15{]}; also see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). These results indicate that in
this data set the fulfillment of Allport's conditions had a significant
influence on outgroup attitudes and this effect is likely, in parts,
explained by its effect through perceived interaction quality.

@@

To test whether Allport's conditions or the core need fulfillment were
better at predicting outgroup attitudes, we first assessed relative
model performance indices (i.e., Akaike information criterion, and
Bayesian information criterion), and then consider the two predictors in
a joint model to see whether the two approaches predicted the same
variance in outgroup attitudes. When comparing the model selection
indices, we found that the fulfillment of the situational core need
indeed performed slightly better than the model using Allport's
conditions (\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59
\(AIC_{Allport}\), and \(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12
\(BIC_{Allport}\)). Additionally, when considering the predictors
jointly, we find that both significantly predict outgroup attitudes with
similar sized regression parameters (random slopes model; Allport's
Conditions: \textit{b} = 0.16, \textit{t}(1,600) = 4.92, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b}
= 0.14, \textit{t}(1,600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}; also see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

@@

We first assessed the stability of our main analyses across the three
studies. Plotting the effect sizes of each parameter of interest in a
forest plot, as well as the average meta-analytic effect, shows that for
the basic contact hypothesis test outgroup contact had a strong and
consistent effect on outgroup attitudes. Interactions with non-outgroup
members consistently had no meaningful effect on outgroup attitudes
(\fgrref{fig:ContactHypothesis}). While this would be expected from the
general intergroup contact literature, this is not a trivial finding.
Being among the first to assess the contact hypothesis using real-life
intensive longitudinal data, we extend cross-sectional findings to
individual-level assessments. When looking at the fulfillment of core
needs during intergroup contacts, we find that the motivational
mechanism is consistently a meaningful predictor of outgroup attitudes
and interaction quality perceptions (see
\fgrref[ A and B]{fig:AllportNeedFulfillment}). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect.

@@

During the main analyses, we have thus far shown that participants held
more positive outgroup attitudes following intergroup contacts and that
a perceived interaction quality was associated with more positive
outgroup attitudes following an intergroup contact. We have, however,
not brought the two elements of the contact hypothesis together in a
single analysis. To jointly test the effects of contact frequency and
average interaction quality, we ran a linear regression model where
average outgroup attitudes were predicted by the number of interactions
and the average interaction quality ratings of all participants across
the three studies. We did so while controlling for the possible effects
of study-specific differences. To include the study-membership as a
control variable, we used the student sample (Study 2) as the reference
group because it was both the largest and the most homogeneous study.
Looking at the overall model, we found that the model predicted 13.89\%
of the variance in average outgroup attitudes (\textit{F}(9, 189) =
3.39, \textit{p} = 0.001, \textit{R\textsuperscript{2}} = 0.14). Looking
at the individual effects, we found that only the number of outgroup
interactions has a clear association with average outgroup attitudes
(\textit{b} = 0.55, \textit{t}(189) = 2.82, \textit{p} = 0.005,
\textit{95\%CI}{[}0.16, 0.93{]}). The average interaction quality
perceptions had a much smaller effect (\textit{b} = 0.29,
\textit{t}(189) = 2.09, \textit{p} = 0.038, \textit{95\%CI}{[}0.02,
0.56{]}), and importantly we found no interaction effect at all. In
short, the effect of interaction frequency did not depend on the average
interaction quality.

@@

One interesting observation was the importance of contact goals specific
to contact through the medical and public health system. This goal type
was partly specific to the medical professionals (e.g., working with
patients, treatment), but also more broadly interactions of newcomers
with the outgroup majority as patients themselves (e.g., therapeutic
goals). Also health, fitness, and personal improvement goals (e.g.,
sports and music) were common goals that participants shared during the
interactions with the majority group members. Another topic less
commonly found within the literature were more practically oriented
needs, where participants sought to share food and cook together, or had
organizational needs (e.g., housing, cleaning, and co-living). In sum,
almost all extracted topics fall into broader or narrower need concepts
that are commonly discussed within the psychological needs literature
\citep[e.g.,][]{Orehek2018a} and offer insight into a core aspect of the
migration experience that has remained broadly under-explored
\citep[][]{Kreienkamp2022d}\footnote{It should be noted that these topic models are not without limitations, especially because they depend on a small set of hyper parameters that determine the characteristics of the embedding, dimension reduction, clustering, and core term extraction. The authors have have also inspected a major subset of the free-text responses manually and are under the impression that the topics described here accurately represent the broader content.}.

>>>>

After receiving ethical approval from the University Masked for Peer
Review, we recruited 23 non-Dutch migrants using the local paid
participant pool. Participants reported on their interactions for at
least 30 days with two daily measures (capturing the morning and
afternoon). With this design, we aimed at getting 50-60 measurements per
participant (\textit{M} = 53.26, \textit{SD} = 16.72, \textit{total N} =
1,225). This is a common number of measurements found in experience
sampling studies and offers sufficient power to model processes within
and between participants \citep[e.g.,][]{AanhetRot2012}. Participants
were compensated for their participation with up to 34 Euros -- each two
Euros for pre- and post-questionnaire and 50 Eurocents for every
experience sampling measurement. The sample consisted of relatively
young, educated, and western migrants from the global north (\(M_{age}\)
= 24.35, \(SD_{age}\) = 4.73, 19 women, 15 students). The sample
accurately describes the largest groups of migrants in the region
\citep[][]{GemeenteGroningen2015}.

@@

To test the prerequisite effect of intergroup contact, every experience
sampling measurement started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (In person interaction for at least 10 minutes)}''.
Our participants recorded between 2--51 interactions with Dutch outgroup
members (3.23--91.07\% of individual experience sampling measurements;
387 of all 1,225 experience sampling
responses)\footnote{Two participants only recorded two experience sampling measurements each and none of these included outgroup contacts. These participants are removed from any analyses that focus on outgroup contacts.}.

@@

Using a multilevel regression, we find that having an outgroup contact
is indeed associated with significantly more positive outgroup attitudes
(\textit{b} = 2.48, \textit{t}(1,200) = 4.37, \textit{p} \textless{}
.001, \textit{95\%CI}{[}1.37, 3.59{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently). Additionally, while multi-level regressions
are generally robust against unequal cell sizes, we correct for
inequalities by using centered predictors and reintroducing the means as
level two predictors (\citealp{Yaremych2021}; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.

@@

The main proposal of our article is that the success of an outgroup
contact might be explained by whether or not the contact fulfilled the
person's core situational need. This should, in turn, be due to a higher
perceived contact quality. We sequentially test whether the fulfillment
of the core need during an interaction is (1) related to more positive
outgroup attitudes, (2) higher perceived contact quality, and (3)
whether the variance explained by the core need is subsumed by the
perceived contact quality if considered jointly. We find that in the
multilevel models, the fulfillment of core situational needs during
outgroup contacts was associated with more positive outgroup attitudes
(random slopes model; \textit{b} = 0.17, \textit{t}(365) = 2.93,
\textit{p} = 0.004, \textit{95\%CI}{[}0.06, 0.29{]}) and also predicted
higher perceived contact quality (random intercept model; \textit{b} =
0.42, \textit{t}(365) = 8.07, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we consider the
influences of core need fulfillment and contact quality on outgroup
attitudes jointly, we find that virtually all variance is explained by
perceived contact quality (random slopes model; \textit{b} = 0.23,
\textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}, also see \fgrref[-A]{fig:MainPaths}). We thus find support for
our hypotheses and can conclude that in this data set the fulfillment of
core situational needs had a significant influence on outgroup
attitudes. Additionally, this effect is likely explained by its effect
through perceived contact quality.

@@

We tested the most general contact hypothesis as we did for Study 1. We
find that having an outgroup interaction is indeed associated with
significantly more positive outgroup attitudes within the participants
(random slopes model; \textit{b} = 2.83, \textit{t}(4,850) = 3.57,
\textit{p} \textless{} .001, \textit{95\%CI}{[}1.28, 4.38{]}), even
after controlling for having an interaction with a non-Dutch person
(which did not relate to outgroup attitudes independently). We again
added the participant means back into the model. We find that in this
data set participant-level outgroup contact proportions were also a
positive predictor of outgroup attitudes (\textit{b} = 26.55,
\textit{t}(110) = 2.90, \textit{p} = 0.004, \textit{95\%CI}{[}8.61,
44.46{]}). The relative number of non-outgroup interactions showed no
such effect (for full results see \tblref{tab:intergroupGeneralTblLong},
\fgrref{fig:ContactHypothesis}, as well as Online Supplementary Material
A). Thus, in our second data set we also find that outgroup contacts
show a positive effect on outgroup attitudes in the moment. We
additionally find an average between-participant effect of the relative
number of interactions participants had.

@@

We again sequentially tested the core need model as we did for Study 1.
We find that in the multilevel models, the fulfillment of core
situational needs during outgroup contacts was associated with more
positive outgroup attitudes (random slopes model; \textit{b} = 0.13,
\textit{t}(826) = 4.18, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.07, 0.19{]}) and also predicted higher perceived
interaction quality (random slopes model; \textit{b} = 0.40,
\textit{t}(826) = 6.46, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.28, 0.52{]}). Additionally, if we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that virtually all variance is explained by
perceived interaction quality (random slopes model; \textit{b} = 0.17,
\textit{t}(825) = 5.56, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.11, 0.23{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.03,
\textit{t}(825) = 1.47, \textit{p} = 0.141, \textit{95\%CI}{[}-0.01,
0.07{]}; also see \fgrref[-B]{fig:MainPaths} and
\tblref{tab:intergroupNeedsTblLong} for full results). These results are
consistent with the results in Study 1. We, thus, find support for our
hypotheses that the fulfillment of core situational needs had a
significant influence on outgroup attitudes and that this effect is
likely explained by its effect through perceived interaction quality.

@@

We recruited 71 international medical students using contacts within the
University Medical School. We specifically targeted non-Dutch students,
who had recently arrived in the Netherlands. Participants reported on
their interactions for at least 30 days with two daily measures
(capturing the morning and afternoon). With this design, we aimed at
getting 50-60 measurements per participant (\textit{M} = 57.85,
\textit{SD} = 20.68, \textit{total N} = 4,107). As with the previous
studies, this offered sufficient power to model processes within
participants. Participants were compensated in the same manner as during
Study 1. The sample consisted of relatively young migrants (\(M_{age}\)
= 22.68, \(SD_{age}\) = 3.10, 59 women). The sample fairly accurately
describes the local population of young international medical
professionals.

@@

In a multilevel regression, we find that having an outgroup interaction
was again associated with significantly more positive outgroup attitudes
within the participants (random slopes model; \textit{b} = 5.57,
\textit{t}(3,834) = 6.52, \textit{p} \textless{} .001,
\textit{95\%CI}{[}3.90, 7.23{]}), even after controlling for having an
interaction with a non-Dutch (which did not relate to outgroup attitudes
independently; for full results see
\tblref{tab:intergroupGeneralTblLong} and
\fgrref{fig:ContactHypothesis}). Thus, in our third data set we find
that the within-person contemporaneous effect of intergroup contact was
consistent across all three studies.

@@

We tested the situational core needs model analogous to the previous
studies. We find that the fulfillment of the core need during outgroup
contacts was associated with more positive outgroup attitudes (random
slopes model; \textit{b} = 0.19, \textit{t}(1,601) = 5.29, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.12, 0.27{]}) and also predicted
higher perceived interaction quality (random slopes model; \textit{b} =
0.43, \textit{t}(1,605) = 8.35, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.33, 0.54{]}). Additionally, once we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that perceived interaction quality is a
substantially stronger predictor (random slopes model; \textit{b} =
0.17, \textit{t}(1,600) = 6.86, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.22{]}) and the unique variance explained by
core need fulfillment was roughly half of its original effect size
(\textit{b} = 0.11, \textit{t}(1,600) = 3.50, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.05, 0.17{]}; also see
\fgrref[-C]{fig:MainPaths} and \tblref{tab:intergroupNeedsTblLong} for
full results). As with the previous two studies, these results indicate
that in this data set outgroup attitudes were significantly predicted by
the fulfillment of core situational needs and the results suggest that
this might be due to higher perceived interaction quality.

@@

We tested the impact of Allport's conditions in the same manner as we
tested our core needs model. In the multilevel models we find that the
fulfillment of Allport's Conditions during outgroup contacts was
associated with more positive outgroup attitudes (random slopes model;
\textit{b} = 0.22, \textit{t}(1,601) = 5.86, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.15, 0.29{]}) and also predicted higher
perceived interaction quality (random slopes model; \textit{b} = 0.65,
\textit{t}(1,605) = 11.82, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.54, 0.76{]}). Moreover, when we considered the
influences of Allport's Conditions and interaction quality on outgroup
attitudes jointly, we found that perceived interaction quality was a
substantially stronger predictor (random slopes model; \textit{b} =
0.17, \textit{t}(1,600) = 6.25, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.23{]}) and the unique variance explained by
Allport's Conditions was less than half of its original effect size
(\textit{b} = 0.09, \textit{t}(1,600) = 2.89, \textit{p} = 0.004,
\textit{95\%CI}{[}0.03, 0.15{]}; also see
\tblref{tab:intergroupNeedsTblLong}). These results indicate that in
this data set the fulfillment of Allport's conditions had a significant
influence on outgroup attitudes and this effect is likely, in parts,
explained by its effect through perceived interaction quality.

@@

To test whether Allport's conditions or the core need fulfillment were
better at predicting outgroup attitudes, we first assessed relative
model performance indices (i.e., Akaike information criterion, and
Bayesian information criterion), and then consider the two predictors in
a joint model to see whether the two approaches predicted the same
variance in outgroup attitudes. When comparing the model selection
indices, we found that the fulfillment of the situational core need
indeed performed slightly better than the model using Allport's
conditions (\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59
\(AIC_{Allport}\), and \(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12
\(BIC_{Allport}\)). Additionally, when considering the predictors
jointly, we find that both significantly predict outgroup attitudes with
similar sized regression parameters (random slopes model; Allport's
Conditions: \textit{b} = 0.16, \textit{t}(1,600) = 4.92, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b}
= 0.14, \textit{t}(1,600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}; also see
\tblref{tab:intergroupNeedsTblLong}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

@@

We first assessed the stability of our main analyses across the three
studies. Plotting the effect sizes of each parameter of interest in a
forest plot, as well as the average meta-analytic effect, shows that for
the basic contact hypothesis test outgroup contact had a strong and
consistent effect on outgroup attitudes. Interactions with non-outgroup
members consistently had no meaningful effect on outgroup attitudes
(\fgrref{fig:ContactHypothesis}). While this would be expected from the
general intergroup contact literature, this is not a trivial finding.
Being among the first to assess the contact hypothesis using real-life
intensive longitudinal data, we extend cross-sectional findings to
individual-level assessments. When looking at the fulfillment of core
needs during intergroup contacts, we find that the motivational
mechanism is consistently a meaningful predictor of outgroup attitudes
and interaction quality perceptions (see
\fgrref[ A and B]{fig:AllportNeedFulfillment}). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect (see
\fgrref[-C]{fig:AllportNeedFulfillment}).

@@

During the main analyses, we have thus far shown that participants held
more positive outgroup attitudes following intergroup contacts and that
perceived interaction quality was associated with more positive outgroup
attitudes following an intergroup contact. We have, however, not brought
the two elements of the contact hypothesis together in a single
analysis. To jointly test the effects of contact frequency and average
interaction quality, we ran a linear regression model where average
outgroup attitudes were predicted by the number of interactions and the
average interaction quality ratings of all participants across the three
studies. We did so while controlling for the possible effects of
study-specific differences. To include the study-membership as a control
variable, we used the student sample (Study 2) as the reference group
because it was both the largest and the most homogeneous study. Looking
at the overall model, we found that the model predicted 13.89\% of the
variance in average outgroup attitudes (\textit{F}(9, 189) = 3.39,
\textit{p} = 0.001, \textit{R\textsuperscript{2}} = 0.14). Looking at
the individual effects, we found that only the number of outgroup
interactions has a clear association with average outgroup attitudes
(\textit{b} = 0.55, \textit{t}(189) = 2.82, \textit{p} = 0.005,
\textit{95\%CI}{[}0.16, 0.93{]}). The average interaction quality
perceptions had a much smaller effect (\textit{b} = 0.29,
\textit{t}(189) = 2.09, \textit{p} = 0.038, \textit{95\%CI}{[}0.02,
0.56{]}), and importantly we found no interaction effect at all. In
short, the effect of interaction frequency did not depend on the average
interaction quality.

@@

One interesting observation was the importance of contact goals specific
to contact through the medical and public health system. This goal type
was partly specific to the medical professionals (e.g., working with
patients, treatment), but also more broadly highlighted interactions of
newcomers with the outgroup majority as patients themselves (e.g.,
therapeutic goals). Also health, fitness, and personal improvement goals
(e.g., sports and music) were common goals that participants shared
during the interactions with the majority group members. Another topic
less commonly found within the literature were more practically oriented
needs, where participants sought to share food and cook together, or had
organizational needs (e.g., housing, cleaning, and co-living). In sum,
almost all extracted topics fall into broader or narrower need concepts
that are commonly discussed within the psychological needs literature
\citep[e.g.,][]{Orehek2018a} and offer insight into a core aspect of the
migration experience that has remained broadly under-explored
\citep[][]{Kreienkamp2022d}\footnote{It should be noted that these topic models are not without limitations, especially because they depend on a small set of hyperparameters that determine the characteristics of the embedding, dimension reduction, clustering, and core term extraction. The authors have have also inspected a major subset of the free-text responses manually and are under the impression that the topics described here accurately represent the broader content.}.
